---
title: "Brownlows Bets"
author: "Skye Strong"
date: "19 April 2018"
output: html_document
---

<!-- We recommend you use git branches -- create a separate branch for each major section i.e. Introduction, Methods, Results, Inferential Results, and Discussion and Conclusions - leave the YAML block untouched until the final merge where you put in your own name !-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(mosaic)
```


# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

 At the first day of class there are 35 students in Math Statistics class and professor Brownlow always proposes this bet to his students "I will wager $2 to your bet of $1 that at least two people in today's class have  the same birthday.  He tells me he has made that bet 1137 times so far. Where do you think his profits are? Discuss this in a markdown document

# Methods

<!--Decide on your methods:  In this section explain the approach you will take to attack the issue --- how will you decide on a likely estimate of professor Brownlow's earnings or losses after 1137 bets. Also explain how you construct a 95% confidence interval about your estimate of his winnings !-->

In order to calculate an estimate of Professor Brownlow's earnings, we need to create a simulation of the total wagers Professor Brownlow claims to have made. First, we will make a test class of 35 students and compare their birthdays. We create a sample of 365 possible birthdays, extract a random 35 of them (1 birthday per student), and check if any of those birthdays are the same. We then repeat this test 1137 times to match the amount of times Professor Brownlow made the bet. From each test, if any birthdays are duplicated, we add 1 dollar to Brownlow's earnings; if there are no duplicated birthdays, we deduct 2 dollars from Brownlow's earnings. In order to make a more accurate prediction, we will repeat our set of 1137 wagers 100 times and save the resultant earnings in an array.  

From the stored earnings, we can calculate the mean profit to get an idea of how much money Brownlow might have earned after making this bet 1137 times. In order to construct a 95% confidence interval, we will examine the imperical data from our array of profits, and apply the standard deviation to the mean. Going two standard deviations below the mean and two standard deviations above the mean will give the required confidence interval.


#Profit Estimate

<!--Explain and display the results of your method of coming up with the estimate of professor Brownlows betting!-->


We have made a sample of 35 students, tested if any of them share a birthday, and doled out profits depending if Brownlow wins or loses the bet. Next, we repeat that test a total of 1137 times to reflect the potential earnings Brownlow has made in total. Shown as our "brownProfit," the total profit was $582. 
```{r}
betarray <- numeric()
for (j in 1:100){
brownProfit = 0
bdays <- (1:365)
for (i in 1:1137){
students <- sample(bdays, 35, replace = TRUE)
xx <- duplicated(students)
WinLose <- sum(xx)
if (WinLose >= 1) brownProfit = brownProfit + 1 else brownProfit = brownProfit - 2
}
betarray[j]=brownProfit
}
hist(betarray)
favstats(betarray)
```
That loop of total tests is repeated 100 times with profits assigned to an array called "betarray." We then performed the `favstats` function on our array and found that the mean profit was $`r round(mean(betarray))`. In other words, after making his birthday wager 1137 times, and then repeating that 100 more times for accuracy, we can estimate that the total profit Professor Brownlow made was $`r round(mean(betarray))`. 


# Confidence interval of Brownlows earning

<!--Explain and display your method of coming up with a 95% confidence interval of  professor Brownlows-->


We now want to construct a 95% confidence interval. In order to find this range, we will calculate two standard deviations below and above the mean.   
```{r}
est <- mean(betarray)
sd <- sd(betarray)
conf <- c(est - 2*sd, est + 2*sd)
conf
```
Our 95% confidence interval is `r round(conf)`, meaning that if we repeated our test, the mean profit would lie somewhere between this range 95% of the time.


# Discussion and Conclusion

<!--Discussion and conclusion here.  If professor Brownlow is consistantly making or losing money how could he change his payout so that his profits center on zero?  Also discuss what size his class should be in order to have his current scheme balance at zero profit.!-->

In conclusion, by repeating Brownlow's 1137 wagers 100 times using a random sample, we estimated an average profit of $`r round(mean(betarray))`. It is important to note that the number of students in the class, as well as the amount Brownlow wagered, influenced the possible amount of earnings. Below we take a look at how changing either of these variables could change our results.

### Repeated Test - Smaller Class Size  

With a class size of 35 students, Professor Brownlow is making a profit. If there were less students, the chances of them having the same birthday (out of 365 possibilities) decreases, thus Brownlow's chances of making a profit decreases. On the other hand if there were more students, the chances of any two of them having the same birthday would increase, thereby increasing the chances of Brownlow making an even larger profit.  
```{r}
test2 <- numeric()
for (j in 1:100){
brownProfit2 = 0
bdays2 <- (1:365)
for (i in 1:1137){
students2 <- sample(bdays2, 28, replace = TRUE)
xx2 <- duplicated(students2)
WinLose2 <- sum(xx2)
if (WinLose2 >= 1) brownProfit2 = brownProfit2 + 1 else brownProfit2 = brownProfit2 - 2
}
test2[j]=brownProfit2
}
favstats(test2)
```
We see from the above test that just by reducing the class size from 35 students to 28 students, Brownlow's profit changes to loss, with an average $`r round(mean(test2))`. Although the results are not shown, repeating this test with 27 students led to an even larger loss while having 29 students resulted in a small profit. If we factor in the standard deviation, it is possible to conclude that in order for his profit to balance at zero, his class size could be around 28 students.

### Repeated Test - Larger Wager

Brownlow's initial wager of $2 to every student's $1 also affects the potential earnings from the bet. Let's take a look at how increasing the wager (keeping the original 35 students) changes our results.  
```{r}
test3 <- numeric()
for (j in 1:100){
brownProfit3 = 0
bdays3 <- (1:365)
for (i in 1:1137){
students3 <- sample(bdays3, 35, replace = TRUE)
xx3 <- duplicated(students3)
WinLose3 <- sum(xx3)
if (WinLose3 >= 1) brownProfit3 = brownProfit3 + 1 else brownProfit3 = brownProfit3 - 4
}
test3[j]=brownProfit3
}
favstats(test3)
```
By simply doubling Brownlow's original wager of $2 to $4, his mean earnings are down to under $100. Even though one might have assumed that a larger wager would result in a bigger payoff, it appears that the number of times he lost the bet led to a decrease in his profit. Logically, we might conclude that if the wager was increased even more, Brownlow's profit would reduce to zero, and he would end up losing money.
